Allocation Removal by Partial Evaluation in a Tracing JIT

### 2011-01-21 talk/pepm2011/presentation/talk.tex
tweaks
todos left
- add 17 new diagrams that work out the example in detail - add a diagram for the benchmarks

@@

\begin{frame}
  \frametitle{Contribution of our Paper}
  \begin{itemize}
      \item a simple, efficient and effective optimization of heap operations in a trace
      \item using online partial evaluation
      \item fully implemented and in use in large-scale interpreters
  \end{itemize}
  \pause
  \begin{block}{Ingredients}
      \begin{itemize}
          \item a slightly simplified model for objects on the heap
          \item operational semantics of trace operations that manipulate the heap
          \item optimization rules for those operations
      \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}[plain]
  \frametitle{Operations: New}
  \includegraphics[scale=0.8]{figures/new01}
\end{frame}

@@

\begin{frame}
  \frametitle{Optimization by Online Partial Evaluation}
  \begin{itemize}
      \item Trace is optimized using online partial evaluation
      \item part of the runtime heap is modelled in the \emph{static heap}
      \item static heap contains objects that are allocated within the trace
      \item (as opposed to before the trace is executed)
      \pause
      \item operations acting on the static heap can be removed
      \item all others need to be residualized
      \item all fairly straightforward
  \end{itemize}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Optimized Example Trace}
  Trace of \texttt{x = a + b; y = x + c}:
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
i1 = get(a, intval)
i2 = get(b, intval)
i3 = int_add(i1, i2)
\sout{x = new(Integer)}
\sout{set(x, intval, i3)}
\sout{guard_class(x, Integer)}
guard_class(c, Integer)
\sout{i4 = get(x, intval)}
i5 = get(c, intval)
i6 = int_add(\emph{i3}, i5)
y = new(Integer)
set(y, intval, i6)
\end{alltt}
\end{frame}

@@

\begin{frame}
  \frametitle{Benchmark Results}
  \begin{itemize}
      \item to evaluate the optimization we used PyPy's Python interpreter with real-world programs
      \item optimization can remove
      \begin{itemize}
          \item 70\% of all \texttt{new} operations
          \item 14\% of all \texttt{get/set} operations
          \item 93\% of all \texttt{guard} operations
      \pause
      \item Timings improve by a factor between 1.1 and 6.95
      \item outperforming standard Python on all benchmarks but one
      \pause
      \item more details in the paper
      \end{itemize}
  \end{itemize}
\end{frame}

>>>>

% XXX todos:
% note that two fields is a simplification
% have a diagram for the example trace
% lifting can never produce new operations
% have some sort of overview or position indicator for the many graphical slides, too easy to get lost
% more details about benchmarks, diagram?

@@

\begin{frame}
  \frametitle{Contribution of our Paper}
  \begin{itemize}
      \item a simple, efficient and effective optimization of heap operations in a trace
      \item using online partial evaluation
      \item fully implemented and in use in large-scale interpreters
  \end{itemize}
  \pause
  \begin{block}{Ingredients}
      \begin{itemize}
          \item a slightly simplified model for objects on the heap
          \item operational semantics of trace operations that manipulate the heap
          \item optimization rules for those operations, following the operational semantics
      \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}[plain]
  \frametitle{Operations: New}
%  \begin{columns}[c]
%  \begin{column}{10cm}
    \includegraphics[scale=0.8]{figures/new01}
%  \end{column}
%  \begin{column}{2cm}
%  {\tiny 
%  \begin{itemize}
%    \item new
%    \item get
%    \item set
%    \item guard
%  \end{itemize}
%  }
%  \end{column}
%  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Optimization by Online Partial Evaluation}
  \begin{itemize}
      \item Trace is optimized using online partial evaluation
      \item part of the runtime heap is modelled in the \emph{static heap}
      \item static heap contains objects that are allocated within the trace
      \item (as opposed to before the trace is executed)
      \pause
      \item operations acting on the static heap can be executed
      \item follows operational semantics
      \item all others need to be residualized
      \item all fairly straightforward
  \end{itemize}
\end{frame}

@@

\begin{frame}[plain]
    \frametitle{Optimizing the Example Trace}
\only<1>
{\includegraphics[scale=0.8]{figures/ex00}}
\only<2>
{\includegraphics[scale=0.8]{figures/ex01}}
\only<3>
{\includegraphics[scale=0.8]{figures/ex02}}
\only<4>
{\includegraphics[scale=0.8]{figures/ex03}}
\only<5>
{\includegraphics[scale=0.8]{figures/ex04}}
\only<6>
{\includegraphics[scale=0.8]{figures/ex05}}
\only<7>
{\includegraphics[scale=0.8]{figures/ex06}}
\only<8>
{\includegraphics[scale=0.8]{figures/ex07}}
\only<9>
{\includegraphics[scale=0.8]{figures/ex08}}
\only<10>
{\includegraphics[scale=0.8]{figures/ex09}}
\only<11>
{\includegraphics[scale=0.8]{figures/ex10}}
\only<12>
{\includegraphics[scale=0.8]{figures/ex11}}
\only<13>
{\includegraphics[scale=0.8]{figures/ex12}}
\only<14>
{\includegraphics[scale=0.8]{figures/ex13}}
\only<15>
{\includegraphics[scale=0.8]{figures/ex14}}
\only<16>
{\includegraphics[scale=0.8]{figures/ex15}}
\only<17>
{\includegraphics[scale=0.8]{figures/ex16}}
\end{frame}

@@

\begin{frame}
  \frametitle{Benchmark Results}
  \begin{itemize}
      \item to evaluate the optimization we used PyPy's Python interpreter with real-world programs
      \item optimization can remove
      \begin{itemize}
          \item 70\% of all \texttt{new} operations
          \item 14\% of all \texttt{get/set} operations
          \item 93\% of all \texttt{guard} operations
      \pause
      \item Timings improve by a factor between 1.1 and 6.95
      \item outperforming standard Python on all benchmarks but two
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Benchmark}
  \includegraphics[scale=0.4]{figures/benchmarks}
\end{frame}
